# Reproducible Research: Peer Assessment 1

```{r setup}
library(knitr)
#only for local use: @TODO: remove
setwd('/Users/kgeusebroek/dev/xebia/coursera/repdata-004/assignments/week1/RepData_PeerAssessment1')

```

## Loading and preprocessing the data
```{r loading, results="hide"}
dataFile <- unz(normalizePath('activity.zip'), 'activity.csv')
df <- read.csv(dataFile, header=T, stringsAsFactors = FALSE, sep=',')
colnames(df)[1] <- "steps"
colnames(df)[2] <- "date"
colnames(df)[3] <- "interval"
```

## What is mean total number of steps taken per day?

```{r}
aggregated_steps = aggregate( steps ~ date, data=df, FUN=sum )

library(ggplot2)
p = ggplot()
p + geom_histogram(data=aggregated_steps, aes(x=steps))

steps_mean <- mean(aggregated_steps$steps)
steps_median <- median(aggregated_steps$steps)
```

Mean is `r steps_mean` and Median is `r steps_median`


## What is the average daily activity pattern?

```{r}
aggregated_activity = aggregate( steps ~ interval, data=df, FUN=sum )
aggregated_activity$intervalDT <- as.POSIXct(strptime(sprintf("%04s", aggregated_activity$interval), format="%H%M"))

library(ggplot2)
library(scales)
p = ggplot(data=aggregated_activity, aes(x=intervalDT, y=steps, group=1))
p + geom_line() + scale_x_datetime(labels = date_format("%H:%M"), breaks = date_breaks("2 hours"))

max_steps <- max(df[, 1], na.rm =T)
interval_with_max_steps <- df[which(df[, 1] == max_steps), ]
```

The interval with the most steps (`r max_steps`) is: `r interval_with_max_steps$interval` on `r interval_with_max_steps$date`

## Imputing missing values
First count the number of rows containing NA values

```{r}
numberNAs <- sum(is.na(df$steps) | is.na(df$interval) | is.na(df$date))
```
`r numberNAs` rows with NA values

```{r}
mean_steps_by_interval = aggregate( steps ~ interval, data=df, FUN=mean )
merged <- merge(df, mean_steps_by_interval, by.x = "interval", by.y="interval", all.x = TRUE)

df_without_na <- df
df_without_na$steps <- ifelse(is.na(merged$steps.x), merged$steps.y, merged$steps.x)

aggregated_steps = aggregate( steps ~ date, data=df_without_na, FUN=sum )

library(ggplot2)
p = ggplot()
p + geom_histogram(data=aggregated_steps, aes(x=steps))

steps_mean <- mean(aggregated_steps$steps)
steps_median <- median(aggregated_steps$steps)
```

Mean is `r steps_mean` and Median is `r steps_median`


## Are there differences in activity patterns between weekdays and weekends?
```{r}
df_without_na$dateConverted <- as.Date(as.character(df_without_na$date), format="%Y-%m-%d")

df_without_na$wd <- weekdays(df_without_na$dateConverted)
df_without_na$weekend <- ifelse(df_without_na$wd %in% c("Saturday","Sunday"), "weekend", "weekday")

aggregated_activity = aggregate( steps ~ interval + weekend, data=df_without_na, FUN=sum )
aggregated_activity$intervalDT <- as.POSIXct(strptime(sprintf("%04s", aggregated_activity$interval), format="%H%M"))

library(lattice)
with(aggregated_activity, xyplot(steps ~ interval | weekend, type='l', as.table = T))
```
